---
title: "Testing Raser"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

# Global adm. boundaries
[Using the package](https://www.gis-blog.com/r-raster-data-acquisition/) `raster` to get admin. boundaries data for
Austria.

```{r Setup}
#install.packages("raster")
library(raster)
library(tidyverse)
```
-   installs and loads the package

```{r getData}
austria0 <- getData('GADM', country='AUT', level=0)
```
-   Arg_1 specifies data set
    -   GDAM -- global administrative boundaries
-   Arg_2 specifies country
    -   Use ISO A3 country code
-   Arg_3 specifies level of administrative subdivision
    -   0 = country, 1 = state
-   Above code returns country boundaries of Austria

```{r Comparing}
#Get Data
austria0 <- getData('GADM' , country="AUT", level=0)
austria1 <- getData('GADM' , country="AUT", level=1)

#Plot
par(mfcol = c(2,1))
plot(austria0, main="Adm. Boundaries Austria Level 0")
plot(austria1, main="Adm. Boundaries Austria Level 1")
```
-   Code prints Country and State borders for Austria
-   mfrow fixed on pg 4 of (this pdf)[https://seananderson.ca/courses/11-multipanel/multipanel.pdf]

# World Climate
-   Similar process for world climate
    -   Three arguments
        -   Dataset (here, 'worldclim')
        -   Variable (here, 'bio', but also ‘tmin’, ‘tmax’, ‘prec’)
        -   Resolutaion (0.5, 2.5, 5, and 10 (minutes of a degree).)
            -   res=0.5 requires lon + lat
            
```{r Climate}
climate <- getData('worldclim', var='bio', res=2.5)
```

THIS RETURNS A LOT

-   BIO1 = Annual Mean Temperature
-   BIO2 = Mean Diurnal Range (Mean of monthly (max temp – min temp))
-   BIO3 = Isothermality (BIO2/BIO7) (* 100)
-   BIO4 = Temperature Seasonality (standard deviation *100)
-   BIO5 = Max Temperature of Warmest Month
-   BIO6 = Min Temperature of Coldest Month
-   BIO7 = Temperature Annual Range (BIO5-BIO6)
-   BIO8 = Mean Temperature of Wettest Quarter
-   BIO9 = Mean Temperature of Driest Quarter
-   BIO10 = Mean Temperature of Warmest Quarter
-   BIO11 = Mean Temperature of Coldest Quarter
-   BIO12 = Annual Precipitation
-   BIO13 = Precipitation of Wettest Month
-   BIO14 = Precipitation of Driest Month
-   BIO15 = Precipitation Seasonality (Coefficient of Variation)
-   BIO16 = Precipitation of Wettest Quarter
-   BIO17 = Precipitation of Driest Quarter
-   BIO18 = Precipitation of Warmest Quarter
-   BIO19 = Precipitation of Coldest Quarter

Tutorial goes on to plot the annual mean temp
```{r Mean Temp}
#Plot
plot(climate$bio1, main="Annual Mean Temperature")
```

Lets sidebar and do `BIO5 = Max Temperature of Warmest Month`

```{r Max Temp}
#Plot
plot(climate$bio5, main="Max Temperature of Warmest Month")
```

And BIO6 = Min Temperature of Coldest Month
```{r Min Temp}
#Plot
plot(climate$bio6, main="Min Temperature of Coldest Month")
```
# SRTM 90 Elevation
```{r SRTM}
srtm <- getData('SRTM', lon=16, lat=48)
```
-   Arg_1: returns SRTM 90 data
-   Lon: longitude
-   Lat: latitude

```{r Elev1}
plot(srtm)
plot(austria0, add=TRUE)
```

LMAO it's missing a bit of info. Need to tile pieces together.

```{r Mosaic}
# Download two more tiles
srtm2 <- getData('SRTM', lon=13, lat=48)
srtm3 <- getData('SRTM', lon=9, lat=48)

#Mosaic/merge srtm tiles
srtmmosaic <- mosaic(srtm, srtm2, srtm3, fun=mean)

plot(srtmmosaic, main="Elevation (SRTM)")
plot(austria0, add=TRUE)
```


# Testing with Greece
```{r Greece}
#Get Data
greece0 <- getData('GADM' , country="GRC", level=0)
greece1 <- getData('GADM' , country="GRC", level=1)

#Plot
#par(mfrow(2,1))s
plot(greece0, main="Adm. Boundaries Greece Level 0")
plot(greece1, main="Adm. Boundaries Greece Level 1")
```




```{r Test}
# Download two more tiles
#srtmGreece0 <- getData('SRTM', lon=23, lat=43)
srtmGreece1 <- getData('SRTM', lon=23, lat=40)
srtmGreece2 <- getData('SRTM', lon=25, lat=40)
srtmGreece3 <- getData('SRTM', lon=25, lat=37)
srtmGreece4 <- getData('SRTM', lon=23, lat=37)
#srtm3 <- getData('SRTM', lon=9, lat=48)

GreeceMosaic <- mosaic(#srtmGreece0,
                       srtmGreece1,
                       srtmGreece2,
                       srtmGreece3,
                       srtmGreece4,
                       fun=mean)

#plot(srtmmosaic, main="Elevation (SRTM)")
plot(GreeceMosaic, main = "Elevation (SRTM)")

```

```{r}
Elev <- as.data.frame(GreeceMosaic, xy=T)%>%drop_na()
head(Elev)
```
```{r}
#isobands(x=Elev$x,
#         y=Elev$y,
#         z=Elev$layer,
#         levels_low=0,
#         levels_high=3000)

ggplot()+
  geom_tile(data=Elev, aes(x=x,y=y,fill=layer))+
  scale_color_gradient()+
  theme_bw()
```

```{r doesnt work}
Elev2 <- as.raster(srtmGreece1)

isobands(x=Elev2[1,], y=Elev2[2,], z=Elev2[3,], levels_low = 0, levels_high = 3000)
```

```{r dialing in} 
plot(srtmGreece1, main = "Elevation (SRTM)")
```

```{r southern greece}
SGreece <- as.data.frame(srtmGreece1, xy=T, returnclass="sf") %>% drop_na()
head(SGreece)
```

```{r}
ggplot(data=SGreece)+
  geom_tile(aes(x=x,y=y,fill=SGreece[,3]))+
  scale_color_gradient()+
  geom_sf(fill="transparent", data = SGreece)+ #fixes skew???
  theme_bw()
``