\documentclass{turabian-researchpaper}
\input{Assets/Preamble/Preamble}

\addbibresource{Assets/Bibliography/Initial.bib}
\usepackage{gfsdidot}

\begin{document}
\maketitle

I propose to study a corpus of inscriptions from seven towns near the borders of Attica. This corpus will include inscriptions from seven cities: Megara, Pagæ, and Ægosthena; Oropus and Tanagra; Eleusis, and Athens. This thesis will be useful both because of the mapped isogloss lines it will argue and as a display of the programming methods used to parse the large volume of data efficiently. This thesis builds on such scholarship as that of A.-F. Christidis on the history and dialectology of ancient Greek in \textit{A History of Ancient Greek from the Beginnings to Late Antiquity} and that of Nikolaos Papzarkadas on the Bœotian inscriptions in \textit{The Epigraphy and History of Boeotia: New Finds, New Prospects}. An example of a similar methodology to mine is found in \textit{Like Two Pis in a Pod: Author Similarity Across Time in the Ancient Greek Corpus} by Grant Storey and Dr. David Mimno. The article focuses on the works of ancient literary authors rather than inscriptions, but the methods involved in text mining and data management are comparable. The ultimate goals of this thesis will be to develop a limited isogloss map of central Greece and to demonstrate an analytical method that one might expand to studies of larger inscriptional corpora.

This thesis will begin with a description of Attic, Bœotian, and Doric Greek. The description will cover the relationship between the dialects and provide a phonological overview of each to enable a reader to compare them against each other. The thesis will then address the methodology used in the study, focusing on a program written to manage the data. The program will be responsible for scraping the information from online publications, cleaning the data of any formatting and editorial marks, and presenting evidence. The study will treat inscriptions from Attica as the ``standards'' against which to compare those of Megaris and Bœotia.

The next chapter will cover isoglosses not attested in Attica. The data will be organized by region, then by city, and further by century. The third chapter will cover isoglosses shared in some capacity with Attica and will follow the same structure. The thesis will then conclude with a description of the isoglosses organized by century. Finally, the text proper will end with the data used in the study, organized by city then further by century.


\clearpage
\nocite{*}
\printbibliography

\end{document}
